Why shouldn't you use numerical tests to diagnose the linear regression models?

### 2022-07-06 paper_comparison/paper_comparison/paper_comparison.Rmd
update the intro

problem: residual plot diagnostics
conventional test: too sensitive

background:

<!-- week 1 -->
1. residual plot for model diagnostics

a. residual is widely used
b. what are the types of residual plots
c. comparison

<!-- week 2 -->
2. conventional test: F, BP
3. visual test: lineup, theory

<!-- week 4 -->
desc of experiment:
1. simulation setup
2. experimental design
3. result

<!-- week 3 -->
comparison of conventional tests:
1. power (visual test vs. conventional test)
(visual test most different one (everything test, any departure))
plot figure in a paper, desc, exp
2. investigate the difference (gap), give examples
3. conventional is too sensitive
4. make conventional less sensitive (vary alpha)

<!-- last week -->
conclusion:
1. too sensitive, visual test is needed/preferable
2. visual test is infeasible in large scale (expensive)
3. future work (role of computer vision)

@@

Regression diagnostics conventionally involve evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more. Common diagnostic techniques including summary statistics, hypothesis testing, and data plots are essential tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].

## Diagnostic plots

Regression analysis is a field of study with at least a hundred years of history. Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. Regardless of the level of difficulty of the book, one will find the importance and usefulness of diagnostic plots being emphasized again and again. Checking diagnostic plots is also the recommended starting point for validating model assumptions like normality, homoscedasticity and linearity [@anscombe_examination_1963].

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are refereed to as standard residual plots. They are commonly used to identify patterns which are indicative of nonconstant error variance or non-linearity [@cook1982residuals]. Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots. The debt on which type of residuals should be used always present. While raw residuals are the most common output of computer regression software package, by applying a scaling factor, the ability of revealing nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression].

As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project  data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space. Observations with the same projection will be treated as equivalent as they have the same position of the abscissa. Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis, but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis. Hence, in practice, multiple standard residual plots with different horizontal axes will be examined.

<!-- large sample size-->
Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed. Thus, for relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aids to eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.

@@

## Hypothesis testing

In addition to diagnostic plots, researcher may also perform formal tests for detecting model defects. Depends on the alternative, variety of tests can be applied. For example, for testing heteroskedasticity, 
one may use the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979]. And for testing non-linearity, there are RESET test [@ramsey_tests_1969] and F-test.

As discussed in @cook1982residuals, most residual based tests for a particular type of departures from model assumptions are sensitive to other types of departures. Especially, outliers will often  incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers. Furthermore, @montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics. Most importantly, the same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies. But still, the effectiveness of statistical tests shall not be disrespected. Statistical tests have chance to provide analysts with unique information. There are also situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked. One will have no choice but to rely on statistical tests if there is any. A good regression diagnostic practice should be a combination of both methods.

## Visual inference

@@

>>>>

<!-- problem: residual plot diagnostics -->
<!-- conventional test: too sensitive -->

<!-- background:  -->

<!-- week 1 -->
<!-- 1. residual plot for model diagnostics -->

<!-- a. residual is widely used -->
<!-- b. what are the types of residual plots -->
<!-- c. comparison -->

<!-- week 2 -->
<!-- 2. conventional test: F, BP -->
<!-- 3. visual test: lineup, theory -->

<!-- week 4 -->
<!-- desc of experiment: -->
<!-- 1. simulation setup -->
<!-- 2. experimental design -->
<!-- 3. result -->

<!-- week 3 -->
<!-- comparison of conventional tests: -->
<!-- 1. power (visual test vs. conventional test) -->
<!-- (visual test most different one (everything test, any departure)) -->
<!-- plot figure in a paper, desc, exp -->
<!-- 2. investigate the difference (gap), give examples -->
<!-- 3. conventional is too sensitive -->
<!-- 4. make conventional less sensitive (vary alpha) -->

<!-- last week -->
<!-- conclusion: -->
<!-- 1. too sensitive, visual test is needed/preferable -->
<!-- 2. visual test is infeasible in large scale (expensive) -->
<!-- 3. future work (role of computer vision) -->

@@

Regression analysis is a field of study with at least a hundred years of history, and regression diagnostics is one of the essential steps in regression analysis. The diagnostic procedure conventionally involves evaluating the fitness of the proposed model, detecting the presence of influential observations and outliers, checking the validity of model assumptions and many more. In terms of diagnostic techniques, data plots, hypothesis testing, and summary statistics are vital tools for a systematic and detailed examination of the regression model [@mansfield1987diagnostic].

Many of those regression diagnostic methods and procedures are mature and well-established in books first published in the twentieth century, such as @draper_applied_2014, @montgomery_introduction_2012, @belsley_regression_1980, @cook_applied_1999 and @cook1982residuals. Regardless of the level of difficulty, one will find the importance and usefulness of diagnostic plots being emphasized in those books repeatedly. Checking diagnostic plots is also the recommended starting point for validating model assumptions such as normality, homoscedasticity and linearity [@anscombe_examination_1963].

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As suggested by @cook1982residuals, these kinds of diagnostic plots are commonly used to identify patterns that are indicative of nonconstant error variance or non-linearity. Raw residuals and studentized residuals are the two most frequently used residuals in standard residual plots. The debt on which type of residuals should be used is always present. While raw residuals are the most common computer regression software package output, by applying a scaling factor, the ability to reveal nonconstant error variance in standard residual plots will often be enhanced by studentized residuals in small sample size [@gunst2018regression]. As a two-dimensional representation of a model in a $p$-dimensional space, standard residual plots project data points onto the variable of the horizontal axis, which is a vector in $p$-dimensional space. Observations with the same projection will be treated as equivalent as they have the same abscissa. Therefore, standard residual plots are often useful in revealing model inadequacies in the direction of the variable of the horizontal axis but could be inadequate for detecting patterns in other directions, especially in those perpendicular to the variable of the horizontal axis. Hence, in practice, multiple standard residual plots with different horizontal axes will be examined [@cook1982residuals]. Overlapping data points is a general issue in scatter plots not limited to standard residual plots, which often makes plots difficult to interpret because visual patterns are concealed. Thus, for a relatively large sample size, @cleveland1975graphical suggests the use of robust moving statistics as reference lines to give aid to the eye in seeing patterns, which nowadays, are usually replaced with a spline or local polynomial regression line.

@@

In addition to diagnostic plots, analysts may perform formal hypothesis testing for detecting model defects. Depends on the alternative hypothesis that being focused on, variety of tests can be applied. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which both derived from the Lagrange multiplier test (ref here) principle and rely on the asymptotic properties of the null distribution. For testing non-linearity, one may apply the F-test (ref here) to examine the significance of certain polynomial and non-linear forms of the regressors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969].

As discussed in @cook1982residuals, most residual-based tests for a particular type of departures from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is also known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of null hypothesis despite the residuals are well-behaved [@cook_applied_1999]. This can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers. @montgomery_introduction_2012 stated that based on their experience, statistical tests are not widely used in regression diagnostics. The same or even larger amount of information can be provided by diagnostic plots than the corresponding tests in most empirical studies. Not to mention, it is almost impossible to have an exactly correctly specified model in reality. There is a well-known aphorism in statistic stated by George Box - "All models are wrong, but some are useful". This indicates proper hypothesis tests will always reject the null hypothesis as long as the sample size is large enough. The outcome "Not reject" can be interpreted as either "effect size is small" or "sample size is small". The outcome "reject" still doesn't inform us whether and how much the model defects are of actual consequence to the inference and prediction. But still, the effectiveness of statistical tests shall not be disrespected. Statistical tests have chance to provide analysts with unique information. There are situations where no suitable diagnostic plots can be found for a particular violation of the assumptions, or excessive diagnostic plots need to be checked. One will have no choice but to rely on statistical tests if there is any. A good regression diagnostic practice should be a balanced combination of both methods.

@@

# Demographic summary

```{r}
polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(education = fct_relevel(education, "High School or below")) %>%
  ggplot() +
  geom_bar(aes(y = education), orientation = "y") -> education_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(age_group)) -> age_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  mutate(pronoun = fct_relevel(pronoun, "He", "She", "They")) %>%
  ggplot() +
  geom_bar(aes(pronoun)) -> gender_plot

polynomials %>%
  group_by(set) %>%
  summarise(across(age_group:previous_experience, first)) %>%
  ggplot() +
  geom_bar(aes(previous_experience)) -> experience_plot

patchwork::wrap_plots(education_plot, age_plot, gender_plot, experience_plot)
```

<!-- # time taken vs order of the plot -->
<!-- # accuracy vs order of the plot -->
<!-- # num of selection vs order of the plot -->
<!-- # order of the plot? -->
<!-- # summary of the demographic information -->

### 2022-07-06 paper_comparison/paper_comparison/paper_comparison.tex
update the intro

problem: residual plot diagnostics conventional test: too sensitive

background:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  residual plot for model diagnostics
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  residual is widely used
\item
  what are the types of residual plots
\item
  comparison
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  conventional test: F, BP
\item
  visual test: lineup, theory
\end{enumerate}

desc of experiment: 1. simulation setup 2. experimental design 3. result

comparison of conventional tests: 1. power (visual test vs.~conventional
test) (visual test most different one (everything test, any departure))
plot figure in a paper, desc, exp 2. investigate the difference (gap),
give examples 3. conventional is too sensitive 4. make conventional less
sensitive (vary alpha)

conclusion: 1. too sensitive, visual test is needed/preferable 2. visual
test is infeasible in large scale (expensive) 3. future work (role of
computer vision)

@@

Regression diagnostics conventionally involve evaluating the fitness of
the proposed model, detecting the presence of influential observations
and outliers, checking the validity of model assumptions and many more.
Common diagnostic techniques including summary statistics, hypothesis
testing, and data plots are essential tools for a systematic and
detailed examination of the regression model
\citep{mansfield1987diagnostic}.

\hypertarget{diagnostic-plots}{%
\subsection{Diagnostic plots}\label{diagnostic-plots}}

Regression analysis is a field of study with at least a hundred years of
history. Many of those regression diagnostic methods and procedures are
mature and well-established in books first published in the twentieth
century, such as \citet{draper_applied_2014},
\citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
the level of difficulty of the book, one will find the importance and
usefulness of diagnostic plots being emphasized again and again.
Checking diagnostic plots is also the recommended starting point for
validating model assumptions like normality, homoscedasticity and
linearity \citep{anscombe_examination_1963}.

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are refereed to as standard residual plots. They
are commonly used to identify patterns which are indicative of
nonconstant error variance or non-linearity \citep{cook1982residuals}.
Raw residuals and studentized residuals are the two most frequently used
residuals in standard residual plots. The debt on which type of
residuals should be used always present. While raw residuals are the
most common output of computer regression software package, by applying
a scaling factor, the ability of revealing nonconstant error variance in
standard residual plots will often be enhanced by studentized residuals
in small sample size \citep{gunst2018regression}.

As a two-dimensional representation of a model in a \(p\)-dimensional
space, standard residual plots project data points onto the variable of
the horizontal axis, which is a vector in \(p\)-dimensional space.
Observations with the same projection will be treated as equivalent as
they have the same position of the abscissa. Therefore, standard
residual plots are often useful in revealing model inadequacies in the
direction of the variable of the horizontal axis, but could be
inadequate for detecting patterns in other directions, especially in
those perpendicular to the variable of the horizontal axis. Hence, in
practice, multiple standard residual plots with different horizontal
axes will be examined.

Overlapping data points is a general issue in scatter plots not limited
to standard residual plots, which often makes plots difficult to
interpret because visual patterns are concealed. Thus, for relatively
large sample size, \citet{cleveland1975graphical} suggests the use of
robust moving statistics as reference lines to give aids to eye in
seeing patterns, which nowadays, are usually replaced with a spline or
local polynomial regression line.

@@

\hypertarget{hypothesis-testing}{%
\subsection{Hypothesis testing}\label{hypothesis-testing}}

In addition to diagnostic plots, researcher may also perform formal
tests for detecting model defects. Depends on the alternative, variety
of tests can be applied. For example, for testing heteroskedasticity,
one may use the White test
\citep{white_heteroskedasticity-consistent_1980} or the Breusch-Pagan
test \citep{breusch_simple_1979}. And for testing non-linearity, there
are RESET test \citep{ramsey_tests_1969} and F-test.

As discussed in \citet{cook1982residuals}, most residual based tests for
a particular type of departures from model assumptions are sensitive to
other types of departures. Especially, outliers will often incorrectly
trigger the rejection of null hypothesis despite the residuals are
well-behaved \citep{cook_applied_1999}. This can be largely avoided in
diagnostic plots as experienced analysts can evaluate the acceptability
of assumptions flexibly, even in the presence of outliers. Furthermore,
\citet{montgomery_introduction_2012} stated that based on their
experience, statistical tests are not widely used in regression
diagnostics. Most importantly, the same or even larger amount of
information can be provided by diagnostic plots than the corresponding
tests in most empirical studies. But still, the effectiveness of
statistical tests shall not be disrespected. Statistical tests have
chance to provide analysts with unique information. There are also
situations where no suitable diagnostic plots can be found for a
particular violation of the assumptions, or excessive diagnostic plots
need to be checked. One will have no choice but to rely on statistical
tests if there is any. A good regression diagnostic practice should be a
combination of both methods.

\hypertarget{visual-inference}{%
\subsection{Visual inference}\label{visual-inference}}

@@

>>>>

@@

Regression analysis is a field of study with at least a hundred years of
history, and regression diagnostics is one of the essential steps in
regression analysis. The diagnostic procedure conventionally involves
evaluating the fitness of the proposed model, detecting the presence of
influential observations and outliers, checking the validity of model
assumptions and many more. In terms of diagnostic techniques, data
plots, hypothesis testing, and summary statistics are vital tools for a
systematic and detailed examination of the regression model
\citep{mansfield1987diagnostic}.

Many of those regression diagnostic methods and procedures are mature
and well-established in books first published in the twentieth century,
such as \citet{draper_applied_2014},
\citet{montgomery_introduction_2012}, \citet{belsley_regression_1980},
\citet{cook_applied_1999} and \citet{cook1982residuals}. Regardless of
the level of difficulty, one will find the importance and usefulness of
diagnostic plots being emphasized in those books repeatedly. Checking
diagnostic plots is also the recommended starting point for validating
model assumptions such as normality, homoscedasticity and linearity
\citep{anscombe_examination_1963}.

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As suggested by \citet{cook1982residuals},
these kinds of diagnostic plots are commonly used to identify patterns
that are indicative of nonconstant error variance or non-linearity. Raw
residuals and studentized residuals are the two most frequently used
residuals in standard residual plots. The debt on which type of
residuals should be used is always present. While raw residuals are the
most common computer regression software package output, by applying a
scaling factor, the ability to reveal nonconstant error variance in
standard residual plots will often be enhanced by studentized residuals
in small sample size \citep{gunst2018regression}. As a two-dimensional
representation of a model in a \(p\)-dimensional space, standard
residual plots project data points onto the variable of the horizontal
axis, which is a vector in \(p\)-dimensional space. Observations with
the same projection will be treated as equivalent as they have the same
abscissa. Therefore, standard residual plots are often useful in
revealing model inadequacies in the direction of the variable of the
horizontal axis but could be inadequate for detecting patterns in other
directions, especially in those perpendicular to the variable of the
horizontal axis. Hence, in practice, multiple standard residual plots
with different horizontal axes will be examined
\citep{cook1982residuals}. Overlapping data points is a general issue in
scatter plots not limited to standard residual plots, which often makes
plots difficult to interpret because visual patterns are concealed.
Thus, for a relatively large sample size, \citet{cleveland1975graphical}
suggests the use of robust moving statistics as reference lines to give
aid to the eye in seeing patterns, which nowadays, are usually replaced
with a spline or local polynomial regression line.

@@

In addition to diagnostic plots, analysts may perform formal hypothesis
testing for detecting model defects. Depends on the alternative
hypothesis that being focused on, variety of tests can be applied. For
example, the presence of heteroskedasticity can usually be tested by
applying the White test \citep{white_heteroskedasticity-consistent_1980}
or the Breusch-Pagan test \citep{breusch_simple_1979}, which both
derived from the Lagrange multiplier test (ref here) principle and rely
on the asymptotic properties of the null distribution. For testing
non-linearity, one may apply the F-test (ref here) to examine the
significance of certain polynomial and non-linear forms of the
regressors, or the significance of proxy variables as in the Ramsey
Regression Equation Specification Error Test (RESET)
\citep{ramsey_tests_1969}.

As discussed in \citet{cook1982residuals}, most residual-based tests for
a particular type of departures from model assumptions are sensitive to
other types of departures. It is likely the null hypothesis is correctly
rejected but for the wrong reason, which is also known as the ``Type III
error''. Additionally, outliers will often incorrectly trigger the
rejection of null hypothesis despite the residuals are well-behaved
\citep{cook_applied_1999}. This can be largely avoided in diagnostic
plots as experienced analysts can evaluate the acceptability of
assumptions flexibly, even in the presence of outliers.
\citet{montgomery_introduction_2012} stated that based on their
experience, statistical tests are not widely used in regression
diagnostics. The same or even larger amount of information can be
provided by diagnostic plots than the corresponding tests in most
empirical studies. Not to mention, it is almost impossible to have an
exactly correctly specified model in reality. There is a well-known
aphorism in statistic stated by George Box - ``All models are wrong, but
some are useful''. This indicates proper hypothesis tests will always
reject the null hypothesis as long as the sample size is large enough.
The outcome ``Not reject'' can be interpreted as either ``effect size is
small'' or ``sample size is small''. The outcome ``reject'' still
doesn't inform us whether and how much the model defects are of actual
consequence to the inference and prediction. But still, the
effectiveness of statistical tests shall not be disrespected.
Statistical tests have chance to provide analysts with unique
information. There are situations where no suitable diagnostic plots can
be found for a particular violation of the assumptions, or excessive
diagnostic plots need to be checked. One will have no choice but to rely
on statistical tests if there is any. A good regression diagnostic
practice should be a balanced combination of both methods.

@@

\hypertarget{demographic-summary}{%
\section{Demographic summary}\label{demographic-summary}}

\includegraphics{paper_comparison_files/figure-latex/unnamed-chunk-2-1.pdf}
